# (PART) Discussion and conclusions {-}

# Discussion and perspectives {#chap8}

```{r setupCH8, include = FALSE, message = FALSE, warning = FALSE, results = "hide"}
library(tidyverse)
library(knitr)
library(here)

# setting seed for reproducibility
set.seed(666)

# setting up knitr options
opts_chunk$set(
    cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE,
    out.width = "100%", fig.align = "center"
    )
```

\initial{S}everal lines of research have suggested that inner speech may involve speech motor processes. This work includes introspective and phenomenological studies, mental chronometry studies, motor interference studies, modelling work as well as neurophysiological and psychophysiological studies (see our short historical review in Chapter \@ref(intro)). However, the involvement of motor processes during inner speech is highly variable between individuals, tasks, and studies. Therefore, it seems reasonable to assume that inner speech comes in different *varieties* that may involve motor processes to a variable degree. We examined this idea by studying the involvement of the speech motor system during induced rumination, a negative and repetitive form of inner speech. In addition to shedding light upon the nature of inner speech, this work may offer new theoretical and experimental tools to assess the presence and persistence of ruminative thoughts.

<!--
More precisely, we used surface electromyography and articulatory suppression to quantify the involvement of the speech motor system during verbal and non-verbal induced rumination.
-->

## Summary of the results

As argued in Chapter \@ref(intro), the guiding assumption underlying this work was that (verbal) rumination may be considered as a form of inner speech. Honouring that assumption, we studied induced rumination with the tools and methods used to investigate the phenomenon of inner speech. In the first experimental chapter (Chapter \@ref(chap3)), we used surface EMG to assess the predictions of two competing views of inner speech production. According to the *motor simulation view*, inner speech would be similar to overt speech, except that final execution of the speech actions is inhibited. Therefore, it should be possible to record peripheral muscular activation in the speech muscles during inner speech. According to the *abstraction view*, the level of truncation between overt speech and inner speech would be higher in the sense that inner speech would not include articulatory features. Therefore, under this view, it is not expected to record peripheral muscular activation in the speech muscles during inner speech. We observed that the...

## Theoretical implications of the results

### Epistemological interlude

In order to fully apprehend the theoretical implications of these results, it might be useful to first clearly articulate the logical argument elaborated throughout the present work. In the first part (the EMG studies presented in Chapter \@ref(chap3) and \@ref(chap4)), the logical argument was as follows: *if* verbal rumination is a form of inner speech, *then* rumination should be accompanied by peripheral muscular activity in the speech muscles. However, going from the substantive hypothesis (verbal rumination is a form of inner speech) to the experimental prediction (i.e., connecting theory to observations) actually requires the use of *auxiliary hypotheses* or *assumptions*. Elucidating these auxiliary assumptions, the actual logical argument from the first part can be restated as follows^[We recognise that this formulation may still be incomplete as some additional auxiliary or instrumental hypotheses may still be incorporated in order to draw a more exhaustive picture of the argument.]:

- *Theoretical assumption ($T$)*: Verbal rumination is a form of inner speech

  - *Auxiliary hypothesis 1 ($A_{1}$)*: Some forms of inner speech involve the motor simulation of speech production
  - *Auxiliary hypothesis 2 ($A_{2}$)*: The simulation mechanism recruits neural networks engaged in (overt) execution
  - *Auxiliary hypothesis 3 ($A_{3}$)*: The motor commands generated during simulation are only partially inhibited
  - *Instrumental hypothesis 1 ($I_{1}$)*: Surface electromyography is a reliable tool to peripherally record partially inhibited motor commands
  - *Ceteris paribus clause ($C_{p}$)*: We assume there is no other factor exerting an appreciable influence that could obfuscate the main effect of interest

- *Prediction*: Induced rumination should be accompanied by peripheral muscular activity (EMG traces) in the speech muscles

In other words, we say that if the ensemble of premises $p$ (i.e., the conjunction of the theoretical assumption, auxiliary hypotheses, etc.) is true, it should follow that $q$ is true. Therefore, stating $p$ suffices to conclude $q$ (modus ponens), that is, $p$ entails $q$. To be even more precise, when we test a theory predicting that if $O_{1}$ (some experimental manipulation or predictor variable), then $O_{2}$ (some observation or measured variable), what we actually say is that this relation holds *if and only if* all the conjuncts above are true. Thus, the logical structure of an empirical test of a theory can be described as the following conceptual formula [@meehl_appraising_1990;@harlow_problem_1997]:

$$
(T \land A_{1} \land A_2 \land A_{3} \land I_{1} \land C_{p} \land C_{n}) \to (O_{1} \supset O_{2})
$$

where the "$\land$" are conjunctions ("and"), the arrow "$\to$" denotes deduction ("follows that ..."), and the horseshoe "$\supset$" is the material conditional ("If $O_{1}$, Then $O_{2}$"). $A_{t}$ is a conjunction of auxiliary theories, $C_{p}$ is a *ceteribus paribus* clause (i.e., we assume there is no other factor exerting an appreciable influence that could obfuscate the main effect of interest), $I_{1}$ is an auxiliary theory regarding instruments, and $C_{n}$ is a statement about experimentally realised conditions (i.e., we assume that there is no systematic error/noise in the experimental settings).

In other words, we imply that a conjunction of all the elements on the left-side (including our substantive theory $T$) does imply the right side of the arrow, that is, "if $O1$, then $O2$". From there, observing $q$ (where $q$ represents the right-side of the above formula) does not allow inferring $p$ (affirming the consequent fallacy) but not observing $p$ ($\lnot p$) allows inferring not $p$ ($\lnot p$) via the modus tollens. However, not observing $q$ does not permit to refute the substantive hypothesis $T$ alone. Rather, not observing $q$ only allows for the refutation of $p$, the *conjunction* of all elements described above (i.e., $T \land A_{1} \land A_2 \land A_{3} \land I_{1} \land C_{p} \land C_{n}$). Put formally, negating the conjunction is logically equivalent to stating a disjunction of the conjuncts [i.e., either one or the other of the conjuncts is false; @meehl_appraising_1990]. Therefore, not observing $q$ only allows for a refutation of $p$ to an extent that is function of the (im)plausibility of the other conjuncts in $p$ (i.e., $A_{1}$, $A_{2}$, $A_{3}$, $I_{1}$, $C_{p}$ and $C_{n}$). To sum up, failing to observe a predicted outcome does not necessarily mean that the theory itself is wrong, but rather that the conjunction of the theory and the underlying assumptions at hand are invalid [@lakatos_falsification_1976;@meehl_appraising_1990;@harlow_problem_1997].

Similarly, the logical argument from the second part (i.e., the relaxation experiments presented in Chapter \@ref(chap3) and \@ref(chap4) as well as the articulatory suppression studies presented in Chapter \@ref(chap6) and \@ref(chap7)) was of the following form: *if* verbal rumination is a form of inner speech, *then*, a disruption of the speech motor system should disrupt rumination. Again, this argument may be restated in a more detailed form as follows:

- *Theoretical assumption ($T$)*: Verbal rumination is a form of inner speech

  - *Auxiliary hypothesis 1 ($A_{1}$)*: Some forms of inner speech involve the motor simulation of speech production
  - *Auxiliary hypothesis 2 ($A_{2}$)*: The simulation mechanism recruits neural networks engaged in (overt) execution
  - *Ceteris paribus clause ($C_{p}$)*: We assume there is no other factor exerting an appreciable influence that could obfuscate the main effect of interest

- *Prediction*: A disruption of the speech motor system should disrupt rumination

In other words, using the same reasoning as before, we say that not observing $q$ only counts as a refutation of $T$ to an extent that is function of the (im)plausibility of the other conjuncts in $p$. The question remains to know how we could assess the plausibility of each conjunct in order to examine the validity of the substantive hypothesis. Interestingly, @strevens_bayesian_2001 discusses a Bayesian solution to this problem (known as the Duhem-Quine problem in philosophy of science). Reformulating the problem as one of assigning "credit or blame to central hypotheses vs. auxiliary hypotheses" [@gershman_how_2019], Strevens suggests a Bayesian framework for confirmation. Let $h$ denotes the substantive hypothesis, $a$ denotes the auxiliary hypothesis, and $d$ denote the data. After observing the data $d$, the prior probability of the conjunct $ha$ (i.e., $p(ha)$) is updated to the posterior distribution $p(ha|d)$ according to Bayes' rule:

$$
P(h a | d) = \frac{P(d | h a) P(h a)}{P(d | h a) P(h a)+P(d | \neg(h a)) P(\neg(h a))},
$$

where $p(d|ha)$ is the *likelihood* of the data under $ha$, and $\lnot (ha)$ denotes the negation of $ha$. From there, marginalising over all possible auxiliary hypotheses, the sum rule of probability allows us to obtain the updated belief about the substantive hypothesis:

$$P(h | d) = P(h a | d) + P(h \neg a | d).$$
Similarly, the marginal posterior over the auxiliary is given by:

$$P(a | d) = P(h a | d) + P(\neg h a | d).$$
To sum up, although failing to observe an outcome predicted by a substantive theory cannot count as a strict falsification of that theory, a Bayesian confirmationist framework permits to assess the plausibility of each conjunct separately and to guide the rational updating of knowledge in the light of incoming data [for more details, see @gershman_how_2019;@strevens_bayesian_2001]. In the next section, we revisit our results, keeping these concepts in mind, in order to assess the plausibility of each conjunct and the evolution of these plausibilities throughout the data accumulated in our work.

```{r, eval = FALSE}
# Reproducing Figure 1 from Gershman (2019) and Strevens (2001)

prior <- seq(from = 0, to = 1, length.out = 101)
auxiliary <- seq(from = 0, to = 1, length.out = 101)
grid <- crossing(prior, auxiliary)
grid$posterior <- grid$prior * (1 - grid$auxiliary) / (1 - grid$auxiliary * grid$prior)

grid %>%
  na.omit %>%
  filter(prior %in% c(0.4, 0.6, 0.8) ) %>%
  ggplot(aes(x = auxiliary, y = posterior / prior, colour = as.factor(prior) ) ) +
  #geom_abline(intercept = 1, slope = - 1, linetype = 3, size = 0.5) +
  geom_line(size = 0.75) +
  theme_bw(base_size = 12) +
  labs(x = "p(a|h)", y = "p(h|d)/p(h)", colour = "p(h)")
```

### Re-reading our results

...

### Implication of these results for inner speech theories

...

### Implication of these results for rumination theories

...

<!--

A revised verson of @mackay_constraints_1992 constraints...

Condensation in rumination, look into think-aloud protocols (e.g. Lyubomirsky et al., 1999) ?

Différence entre petit effet homogène (dominant) et > effet hétéogène, réf aux travaux de haaf et rouder, further work should probably explore the domnancy of the effect

From Friston (2011): there is no need for separate inverse and forward models in motor control because the inverse model can be replaced by (Bayesian) inversion of the forward model […] page 491/ Active inference eschews the hard inverse problem by replacing optimal control signals that specify muscle movements (in an intrinsic frame) with prior beliefs about limb trajectories (in an extrinsic frame)  

From Pickering & clarke, IFM replace the need for an inversme model (AFM) ?

See also wlikinson & fernyough (2017)

Computational modelling ? For instance, model of motor imagery duration (in relation to execution duration), maybe find some inspiration in Tversky & Kahneman (1992) model of probability of winning a gamble vs subjective probability ? Cf. description of the model in Farell & Lewandowsky (2018, p. 252-253).

Clarify the logical chain derivation chain from inner speech to EMG correlates (i.e., inner speech = motor simulation = rehearsal of the neuro-motor networks implicated in speech production = incomplete inhibition = EMG traces)

Posterior prob of central hypo and auxiliary hypo ? Computing them via Bayes theorem ? See Gershman (2018; p.15)

How to overcome (bad) mental habits ? From Hertel (2004, page 209): In short, the best antidote to maladaptive habits is a new set of habits—not the opposite sort of habits recommended by Pollyanna and the teachings of Norman Vincent Peale (1956) and not the habits of suppression as recommended by certain grandmothers but the habits of thought control.

Describe model of van vugt... suractivation of neagtive memory chunks ? Alternatively, rumination can be seen in a Bayesian framework as overdetermined priors… or lack of flexibility.

Mental habit = low control (automaticity). Habits become habits through learning and association, in the same way a poem can be learned by heart… both can be more and more strongly internalised in a similar way…

See fitzgerald et al. (2014): how habits are created ? Through model comparison and model averaging, simpler model (habits models) are favoured over time

From Dolan & Doyan (2014): Model-free control is computationally efficient, since it replaces computation (i.e., the burdensome simulation of future states) with memory (i.e., stored discounted values of expected future reward); however, the forward-looking nature of the pre- diction error makes it statistically inefficient (Daw et al., 2005).

>> More automatic inner speech relies more on associate memory-based processes while less automatic (more intentional / deliberate) relies more on deliberate simulation mechanisms. Why ? Because habitual motor action relies on memory-based stored association (mappings) between motor command and sensory consequences. 

In contrast, new (unusual) motor commands, to be imagined, needs to go through the simulation/emulation 
mechanism. This could be tested by creating habits (via learning) of differents words and comparing their EMG traces and / or suppression by articulatory suppression… For instance, mental multiplication ? Learned arithmetic table ? Poems ? Also see: https://link.springer.com/content/pdf/10.3758%2Fs13421-013-0320-y.pdf

The condensation dimension might not be related to “embodiment” (e.g., EMG traces) as automatic thoughts vary along that axis… (e.g., an earworm).

Perspectives: AFM ou IFM ? Ouverture vers post-doc et mécanismes inhibiteurs

-->

## Limitations and ways forward

...

## Conclusion

...
