# (PART) Discussion and conclusion {-}

# Discussion and perspectives {#chap8}

```{r setupCH8, include = FALSE, message = FALSE, warning = FALSE, results = "hide"}
library(tidyverse)
library(knitr)
library(here)

# setting seed for reproducibility
set.seed(666)

# setting up knitr options
opts_chunk$set(
    cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE,
    out.width = "100%", fig.align = "center"
    )
```

\initial{S}everal lines of research have suggested that inner speech may recruit speech motor processes. This work includes introspective and phenomenological studies, mental chronometry studies, motor interference studies, modelling work as well as neurophysiological and psychophysiological studies (cf. our short historical review in Chapter \@ref(intro)). However, the involvement of motor processes during inner speech is highly variable between individuals, tasks, and studies. Therefore, it seems reasonable to assume that inner speech comes in different *varieties* that may involve speech motor processes to a variable extent. We tested this assumption by examining the involvement of the speech motor system during induced rumination, a negative and repetitive form of inner speech. In addition to shedding light upon the nature of inner speech, this work may offer new theoretical and experimental tools to assess the presence and persistence of ruminative thoughts.

## Summary of the results {#summary}

As argued in Chapter \@ref(intro), the guiding assumption underlying this work was that (verbal) rumination may be considered as a form of inner speech. Honouring that assumption, we studied induced rumination with the tools and methods used to investigate the phenomenon of inner speech. In the first experimental chapter (Chapter \@ref(chap3)), we used surface EMG to assess the predictions of two competing views of inner speech production. According to the *motor simulation view*, inner speech would be similar to overt speech, except that final execution of the speech actions is inhibited. Therefore, it should be possible to record peripheral muscular activation in the speech muscles during inner speech. According to the *abstraction view*, the level of truncation between overt speech and inner speech would be higher in the sense that inner speech would not include articulatory features. Therefore, under this view, it is not expected to record peripheral muscular activation in the speech muscles during inner speech. We observed that the induction of rumination was accompanied with increased EMG amplitude in all facial muscles (i.e., orbicularis oris inferior, orbicularis oris superior and frontalis) as compared to rest. It should be noted that baseline recordings were performed after a relaxation session. We interpreted these findings as a corroboration of the motor simulation view, with the activity of the speech muscles increasing from baseline to after the induction, suggesting that rumination, as a form of inner speech, may involve the speech motor system. We interpreted the increased activation of the forehead muscle as a consequence of the negative content of rumination, as the frontalis muscle is known to be associated with the expression of anger and sadness. Additionally, in the second part of this experiment, we observed that a relaxation focused on the speech muscles was slightly more efficient than a relaxation focused on the arm in reducing self-reported state rumination. However, a few important limitations are worth keeping in mind when considering these results. First, as we did not have any control group for the rumination induction, it is adventurous to attribute the observed effects to the rumination induction only.^[More precisely, concluding $p$ on the basis of $q$ would be committing the "affirming the consequent" fallacy, known formally as $\frac{p \rightarrow q, q}{\therefore p}$. In other words, observing $q$ is insufficient to conclude $p$ because $q$ might have been observed for other reasons than $p$. In our situation, the EMG amplitude might have increased for other reasons than the rumination induction. For instance, it might be that the EMG amplitude was higher after induction only because we compared muscular activity after induction to a relaxation period (which might show a lower than usual level of muscular activity). Alternatively, the increase in EMG amplitude might be due to the fact that participants were doing something, in opposition to doing nothing (i.e., this increase may not be specific to rumination).] Second, and related to our first point, the dissociation between the activation of the lip muscles and the forehead muscle may not be that straightforward. Indeed, it might be that the activation of both sites was related to rumination as a form of inner speech, or that the activity of all facial muscles was related to negative affects only. This point echoes and strengthens the recommendations made by @garrity_electromyography_1977 and discussed in Box \ref{garrity}.

To tackle these limitations, in Chapter \@ref(chap4), we sought to examine the differences between different forms of induced rumination, that should theoretically involve the speech motor system to a different extent. We compared the EMG orofacial correlates of either verbal rumination or non-verbal (visual) rumination. Unfortunately, self-reports of the modal content of the ruminative thoughts showed that our induction did not succeed in inducing rumination in different modalities. However, even when exploring the (a posteriori) relation between the modality of the ruminative thoughts and the facial EMG correlates, we failed to find the predicted relation. Put simply, verbal rumination was not associated with more activity in the speech muscles than non-verbal rumination. Moreover, comparing two types of relaxation (as in Chapter \@ref(chap3)) revealed that, in contrast to previous results, the arm relaxation was slightly more efficient than the orofacial relaxation in reducing state rumination. Averaging the relaxation results from these two first studies revealed that both relaxation types have a similar effect on state rumination. These results therefore suggest that verbal rumination is not specifically accompanied with peripheral muscular activity in the speech muscles. However, it is unclear whether this result is due to a poor sensitivity of the surface EMG measurements or to the fact that rumination is a form of inner speech that does not involve the speech motor system.

In Chapter \@ref(chap5), we sought to resolve that ambiguity by examining the reliability of our EMG measurements to detect peripheral muscular activity during inner speech production. To this end, we asked participants to produce two lists of nonwords, that were designed to induce either a strong activation of the lip muscles or a strong activation of the zygomaticus major muscle. We recorded the EMG amplitude of several facial muscles (including the orbicularis oris inferior and the zygomaticus major) during the production of these nonwords in inner speech, overt speech, and during the listening of these nonwords. Based on previous results in the literature, we hypothesised that surface EMG may be used to discriminate the content (here, the class of nonword) produced in inner speech. However, an automatic classification revealed that although we were able to discriminate content produced in overt speech, we were not able to discriminate the content produced in inner speech based on surface EMG measurements. This result stands in contrast with previous historical results but also with more recent results obtained by other teams. However, crucial differences between our study and others include differences in the material used (e.g., surface vs. intramuscular recordings), the population (e.g., children vs. adults), or the general methodology (e.g., hypothesis testing vs. classification and optimisation). Despite this surprising result and the failure of the surface EMG methodology to "decode" the content of inner speech, the abundance of positive results in the literation still speaks in favour of the peripheral muscular components of inner speech and that it is possible to assess it using surface EMG. However, in order to avoid this potential limitation, we shifted in the second part of the present work to another strategy in examining the role of motor processes in rumination. More precisely, instead of recording peripheral muscular activation of induced rumination, we tried to directly interfere with the speech motor system to check whether this would affect verbal rumination.

In Chapter \@ref(chap6), we set up a critical test of the motor simulation view of inner speech (and rumination). Indeed, if the involvement of the speech motor system is necessary during inner speech and rumination, then a disruption of the speech motor system should disrupt (or impair) the production of inner speech (and rumination). To examine this idea, we compared the effects of an articulatory suppression task to a finger-tapping task, following a rumination induction, on the levels of self-reported state rumination. Our results suggest a similar decrease in state rumination following both motor activities, suggesting that rumination would not be a form of inner speech that would crucially depend on the activity of the speech motor system (it does not mean that the speech motor system may never be involved in rumination, only that its involvement is not necessary). However, some important limitations make the interpretation of these results delicate. First, there were important differences between the two groups (i.e., articulatory suppression vs. finger-tapping groups) at baseline, possibly due do the rhythmic training proposed before baseline measurements. Second, the measure of state rumination consisted in one single non-validated scale (already used in Chapter \@ref(chap3)) and may not be a reliable index of state rumination. Third, as in Chapter \@ref(chap3), there was no control group to the rumination induction and any effect following to the rumination induction may not be specifically attributable to the rumination induction. Fourth, there is some evidence suggesting that finger-tapping (with the dominant hand) may also perturbs speech motor planning and may therefore not be the best control condition to articulatory suppression. We sought to overcome these limitations in the last empirical chapter.

Finally, in Chapter \@ref(chap7), we extended the experiment from Chapter \@ref(chap6) by comparing the effects of articulatory suppression (vs. finger-tapping) on rumination and problem-solving, another (more adaptive) form of repetitive thinking. To overcome the limitations of the previous experiment, we used a validated scale of state rumination and asked our participants to use their non-dominant forearm in the finger-tapping condition. We also made sure that our baseline measurements were not contaminated by any systematic effect. The data collection for this experiment is still ongoing and given the very low sample size (around 10 participants per group at the time of writing), we will not consider these results further in the present discussion. However, preliminary analyses presented in the results section of Chapter \@ref(chap7) suggest that articulatory suppression may indeed interfere with induced rumination.

## Theoretical implications of the results

### Epistemological interlude

In order to fully apprehend the theoretical implications of these results, it might be useful to first clearly articulate the logical argument elaborated throughout the present work. In the first part (the EMG studies presented in Chapter \@ref(chap3) and \@ref(chap4)), the logical argument was as follows: *if* verbal rumination is a form of inner speech, *then* rumination should be accompanied by peripheral muscular activity in the speech muscles. However, going from the substantive hypothesis (verbal rumination is a form of inner speech) to the experimental prediction (i.e., connecting theory to observations) actually requires the use of *auxiliary hypotheses* or *assumptions*. Elucidating these auxiliary assumptions, the actual logical argument from the first part can be restated as follows:^[We recognise that this formulation may still be incomplete as some additional auxiliary or instrumental hypotheses may still be incorporated in order to draw a more exhaustive picture of the argument.]

- *Theoretical assumption ($T$)*: Verbal rumination is a form of inner speech

  - *Auxiliary hypothesis 1 ($A_{1}$)*: Some forms of inner speech involve the motor simulation of speech production
  - *Auxiliary hypothesis 2 ($A_{2}$)*: The simulation mechanism recruits neural networks engaged in (overt) execution
  - *Auxiliary hypothesis 3 ($A_{3}$)*: The motor commands generated during simulation are only partially inhibited
  - *Instrumental hypothesis 1 ($I_{1}$)*: Surface electromyography is a reliable tool to peripherally record partially inhibited motor commands
  - *Ceteris paribus clause ($C_{p}$)*: We assume there is no other factor exerting an appreciable influence that could obfuscate the main effect of interest

- *Prediction*: Induced rumination should be accompanied by peripheral muscular activity (EMG traces) in the speech muscles

In other words, we say that if the ensemble of premises $p$ (i.e., the conjunction of the theoretical assumption, auxiliary hypotheses, etc.) is true, it should follow that $q$ is true. Therefore, stating $p$ suffices to conclude $q$ (modus ponens), that is, $p$ entails $q$. To be even more precise, when we test a theory predicting that if $O_{1}$ (some experimental manipulation or predictor variable), then $O_{2}$ (some observation or measured variable), what we actually say is that this relation holds *if and only if* all the conjuncts above are true. Thus, the logical structure of an empirical test of a theory can be described as the following conceptual formula [e.g.,@meehl_appraising_1990;@harlow_problem_1997]:

$$
(T \land A_{1} \land A_2 \land A_{3} \land I_{1} \land C_{p} \land C_{n}) \to (O_{1} \supset O_{2})
$$

where the "$\land$" are conjunctions ("and"), the arrow "$\to$" denotes deduction ("follows that ..."), and the horseshoe "$\supset$" is the material conditional ("If $O_{1}$, Then $O_{2}$"). $A_{t}$ is a conjunction of auxiliary theories, $C_{p}$ is a *ceteribus paribus* clause (i.e., we assume there is no other factor exerting an appreciable influence that could obfuscate the main effect of interest), $I_{1}$ is an auxiliary theory regarding instruments, and $C_{n}$ is a statement about experimentally realised conditions (i.e., we assume that there is no systematic error/noise in the experimental settings).

In other words, we imply that a conjunction of all the elements on the left-side (including our substantive theory $T$) does imply the right side of the arrow, that is, "if $O1$, then $O2$". From there, observing $q$ (where $q$ represents the right-side of the above formula) does not allow inferring $p$ (affirming the consequent fallacy) but not observing $q$ ($\lnot q$) allows inferring not $p$ ($\lnot p$) via the modus tollens. However, not observing $q$ does not permit to refute the substantive hypothesis $T$ alone. Rather, not observing $q$ only allows for the refutation of $p$, the *conjunction* of all elements described above (i.e., $T \land A_{1} \land A_2 \land A_{3} \land I_{1} \land C_{p} \land C_{n}$). Put formally, negating the conjunction is logically equivalent to stating a disjunction of the conjuncts [i.e., either one or the other of the conjuncts is false; @meehl_appraising_1990]. Therefore, not observing $q$ only allows for a refutation of $p$ to an extent that is function of the (im)plausibility of the other conjuncts in $p$ (i.e., $A_{1}$, $A_{2}$, $A_{3}$, $I_{1}$, $C_{p}$ and $C_{n}$). To sum up, failing to observe a predicted outcome does not necessarily mean that the theory itself is wrong, but rather that the conjunction of the theory and the underlying assumptions at hand are invalid [@lakatos_falsification_1976;@meehl_appraising_1990;@harlow_problem_1997].

Similarly, the logical argument from the second part (i.e., the relaxation experiments presented in Chapter \@ref(chap3) and \@ref(chap4) as well as the articulatory suppression studies presented in Chapter \@ref(chap6) and \@ref(chap7)) was of the following form: *if* verbal rumination is a form of inner speech, *then*, a disruption of the speech motor system should disrupt rumination. Again, this argument may be restated in a more detailed form as follows:

- *Theoretical assumption ($T$)*: Verbal rumination is a form of inner speech

  - *Auxiliary hypothesis 1 ($A_{1}$)*: Some forms of inner speech involve the motor simulation of speech production
  - *Auxiliary hypothesis 2 ($A_{2}$)*: The simulation mechanism recruits neural networks engaged in (overt) execution
  - *Ceteris paribus clause ($C_{p}$)*: We assume there is no other factor exerting an appreciable influence that could obfuscate the main effect of interest

- *Prediction*: A disruption of the speech motor system should disrupt rumination

In other words, using the same reasoning as before, we say that not observing $q$ only counts as a refutation of $T$ to an extent that is function of the (im)plausibility of the other conjuncts in $p$. The question remains to know how we could assess the plausibility of each conjunct in order to examine the validity of the substantive hypothesis. Interestingly, @strevens_bayesian_2001 discusses a Bayesian solution to this problem (known as the Duhem-Quine problem in philosophy of science). Reformulating the problem as one of assigning "credit or blame to central hypotheses vs. auxiliary hypotheses" [@gershman_how_2019], Strevens suggests a Bayesian framework for confirmation. Let $h$ denotes the substantive hypothesis, $a$ denotes the auxiliary hypothesis (the reasoning can be generalised to multiple auxiliary hypotheses), and $d$ denote the data. After observing the data $d$, the prior probability of the conjunct $ha$ (i.e., $p(ha)$) is updated to the posterior distribution $p(ha|d)$ according to Bayes' rule:

$$
P(h a | d) = \frac{P(d | h a) P(h a)}{P(d | h a) P(h a)+P(d | \neg(h a)) P(\neg(h a))},
$$

where $p(d|ha)$ is the *likelihood* of the data under $ha$, and $\lnot (ha)$ denotes the negation of $ha$. From there, marginalising over all possible auxiliary hypotheses, the sum rule of probability allows us to obtain the updated belief about the substantive hypothesis:

$$P(h | d) = P(h a | d) + P(h \neg a | d).$$
Similarly, the marginal posterior over the auxiliary is given by:

$$P(a | d) = P(h a | d) + P(\neg h a | d).$$
To sum up, although failing to observe an outcome predicted by a substantive theory cannot count as a strict falsification of that theory, a Bayesian confirmationist framework permits to assess the plausibility of each conjunct separately and to guide the rational updating of knowledge in the light of incoming data [for more details, see @gershman_how_2019;@strevens_bayesian_2001]. In the next section, we revisit our results, keeping these concepts in mind, in order to assess the plausibility of each conjunct and the evolution of these plausibilities throughout the data we accumulated in our work.

```{r, eval = FALSE}
# Reproducing Figure 1 from Gershman (2019) and Strevens (2001)

prior <- seq(from = 0, to = 1, length.out = 101)
auxiliary <- seq(from = 0, to = 1, length.out = 101)
grid <- crossing(prior, auxiliary)
grid$posterior <- grid$prior * (1 - grid$auxiliary) / (1 - grid$auxiliary * grid$prior)

grid %>%
  na.omit %>%
  filter(prior %in% c(0.4, 0.6, 0.8) ) %>%
  ggplot(aes(x = auxiliary, y = posterior / prior, colour = as.factor(prior) ) ) +
  geom_line(size = 0.75) +
  theme_bw(base_size = 12) +
  labs(x = "p(a|h)", y = "p(h|d)/p(h)", colour = "p(h)")
```

### Re-reading our results

What is the role of speech motor processes in rumination? This question can be reframed as follows: what is the role of speech motor processes in inner speech production and how this role varies across the different varieties of inner speech? Are some forms of inner speech always/never motoric? As suggested in the previous discussion, assessing such theoretical issues requires considering and weighting the plausibility of auxiliary assumptions used to connect theoretical statements to empirical predictions. One of the major assumption we have made in the first part of the present work (the EMG studies) was that surface EMG was a reliable way of examining inner speech production. What is the plausibility of that assumption?

As reviewed in Chapter \@ref(intro), many studies have shown that it is possible to use (both surface and intramuscular) electromyography to "decode" the content of inner speech, although some studies failed to do so. However, results from EMG studies of inner speech come in different flavour of persuasiveness depending on the strictness of their experimental protocol. As discussed previously, the most convincing studies are the one showing muscle-specific EMG correlates of inner speech production [e.g.,@mcguigan_patterns_1989;@mcguigan_discriminative_1974]. Most of these studies involve multiple measurements per participant and per stimulus, providing a high sensitivity to these EMG measurements. In contrast to these studies, least convincing studies include the one showing a general (i.e., non-specific) increase in facial muscular activity from rest to the condition of interest, as this increase may be due to many other factors than inner speech production per se (cf. our discussion in the last section but also in Chapter \@ref(intro) and Chapter \@ref(chap5)). Interestingly, the latter type of study usually focuses on more ecological occurrences of inner speech, such as the production of fully formed sentences, poem recitation, or the occurrence of maladaptive forms of inner speech (e.g., AVHs or rumination). One crucial difference between EMG studies of lower-level inner speech studies (like the one we carried out in Chapter \@ref(chap5)) and the EMG studies of more naturalistic forms of inner speech (like the one we carried out in Chapter \@ref(chap3)) is the ability to repeat measurements. Indeed, whereas it is relatively easy to obtain several repetitions of multiple vowels or syllables for a given pool of participant, it is experimentally more arduous to repeat the measurement of more complex forms of inner speech such as AVHs or rumination. As a consequence, and given that the sensitivity of our EMG measures was already insufficient to decode inner speech in the experiment reported in Chapter \@ref(chap5), it might be that the sensitivity of surface EMG was also too low to detect the presence of rumination as a form of inner speech (as observed in Chapter \@ref(chap3) and Chapter \@ref(chap4)). Importantly, we mean that the sensitivity of surface EMG is too low to detect the presence of rumination *as a form of inner speech*. It does not mean that surface EMG can not be used to assess the presence of rumination (e.g., focusing on the activity of the frontalis or any other facial muscle), only that the changes in EMG amplitude can not be attributed to speech motor processes per se (because they are not muscle-specific and not specific to the verbal content of rumination). To sum up, although surface EMG measurements may be used to assess the content of inner speech production (in sufficiently well powered experimental designs), it might not be used to assess the presence of more naturalistic and uniquely (i.e., on a single occasion) occurring forms of inner speech. To put it in other words, although our instrumental assumption $I_{1}$ about the reliability of EMG measurements may be valid in well-powered designs, it may not be valid in EMG studies of rumination (which may impede the ability to test our substantive hypotheses regarding the role of motor processes in inner speech and rumination). We shall now examine the results from the second part of the thesis, in which we directly tried to interfere with the speech motor system during rumination.

Results from this second part include the results from the relaxation experiments reported in Chapter \@ref(chap3) and \@ref(chap4) as well as the articulatory suppression study presented in Chapter \@ref(chap6) (as discussed previously, we will not consider the preliminary results from Chapter \@ref(chap7)). As discussed in section \@ref(summary), the combined results from the relaxation experiments presented in Chapter \@ref(chap3) and \@ref(chap4) suggest that a relaxation focused on the orofacial area was not more efficient than a relaxation focused on the non-orofacial (brachial) area. This observation is interesting in many ways. First, it highlights the need for replication. Indeed, based on the results from Chapter \@ref(chap3) only, we would have concluded that state rumination could be reduced via targeted relaxation. Reciprocally, based on the results from Chapter \@ref(chap4) only, we would have concluded that state rumination could be reduced by relaxation focused on the arm. It is only the combined consideration of both results that allowed us to observe this null effect. Second, it shows that although relaxation may decrease state rumination, the effect of relaxation is not specific to motor processes (both types of relaxation were equally effective, on average, in reducing state rumination). These results suggest that the activity of the speech motor system is not *necessary* for experiencing rumination (it does not mean that the activity of the speech motor system does not play a role at all), as state rumination was not affected by manipulation of the state of the speech motor system. This observation is coherent with previous results, showing that a "passive" peripheral disruption of the speech motor system (e.g., via anaesthesia or relaxation) generally does not disrupt inner speech. In complement to these experiments, we used articulatory suppression in Chapter \@ref(chap6) to directly interfere with the speech motor system during rumination. This operationnalisation may be said to differ from the relaxation experiment in the sense that this requires the participant to actively plan speech motor actions. Therefore, we consider this as an "active" peripheral disruption of the speech motor system. Results from this study are difficult to interpret however, for the reasons already mentioned in Chapter \@ref(chap6) and section \@ref(summary). However, these results suggest that articulatory suppression was not more efficient in reducing state rumination than finger-tapping. This results again corroborates the idea that the activity of the speech motor system is not *necessary* for experiencing rumination (although this conclusion should be further confirmed or contradicted by the data collected in the experiment from Chapter \@ref(chap7)).

Overall, these results suggest that rumination does not *necessitate* the activity of the speech motor system (as it was not associated with specific activity in the speech muscles and as it was not affected by motor interference). In the next sections, we discuss the implications of these results for theories of inner speech and rumination and suggest ways forward from an experimental perspective.

### Implication of these results for inner speech theories

How could it be that some forms of inner speech involve the speech motor system to an extent that is quantifiable using surface EMG whereas some others forms (e.g., rumination) do not? This question can be attacked from different perspectives and at different levels of explanation. As discussed in Chapter \@ref(intro), Vygotsky's model of inner speech development and @fernyhough_alien_2004's extended four-level model suggest that inner speech may be expressed with different degrees of "externalisation", from condensed inner speech to expanded inner speech. These forms of inner speech seat on a continuum and it seems legitimate to assume that more expanded forms of inner speech recruit the speech motor system to a greater extent than more condensed forms of inner speech. This idea is supported by many studies showing a progressive externalisation of inner speech under cognitively demanding situations [e.g.,@sokolov_inner_1972]. However, these models do not stipulate how the involvement of the speech motor system is regulated. What mechanism(s) may explain the differences in the degree of involvement of the speech motor system during inner speech production?

As discussed previously, @sokolov_inner_1972 also observed that the "externalisation" of inner speech^[By "externalisation" we mean here the degree to which inner speech recruits the speech motor system, with fully externalised speech corresponding to ordinary overt speech production.] was a function of the novelty of the task and of the degree of automaticity. How could the difficulty, novelty, and automaticity of the task influence the externalisation of inner speech? We already outlined some possible answers to this question in Chapter \@ref(intro). According to @cohen_motor_1986, the presence of motor activity during inner speech may be interpreted in terms of attentional sharing. For instance, cognitively demanding situations (e.g., novel or difficult tasks) arguably require greater amount of attention to be performed. In these situations, the vividness of inner speech percepts could be strengthened by increasing the speech motor activity. Alternatively, the greater externalisation of inner speech in cognitively demanding tasks may be interpreted in the motor control framework by postulating that lower amount of inhibition will be applied to block motor commands during inner speech, resulting in higher levels of motor activity (and also arguably more vivid inner speech percepts). These two explanations are not incompatible and as discussed previously, the modulation of the amount of inhibition provides a mechanism through which inner speech percepts are reinforced (or not).

How does this fit with our results and with rumination more specifically? As we will argue, rumination can be considered as a *mental habit*, that is, a mental process that became automatic by repetition (cf. our more detailed discussion in the next section). As discussed in section \@ref(motor-imagery), the peripheral muscular activation often observed during motor imagery and inner speech may be attributed to (the consequences of) partially inhibited motor commands (i.e., to residual movements). Therefore, variations in the amount of peripheral muscular activity recorded during inner speech production may be attributed to variation in the amount of inhibition applied to motor commands issued during inner speech production. Thus, rumination, may be considered as a strongly internalised form of inner speech that does not recruit the speech motor system (in other words, the motor commands that emitted during this form of inner speech are greatly inhibited). However, it is still unclear what exactly these inhibitory mechanisms are [e.g., what kind of inhibition are we talking about, @gorfein_concept_2007; how and when are these mechanisms are implemented, cf. @guillot_imagining_2012] and elucidating the inhibitory mechanisms underlying inner speech production will be the focus of one of our future research project.

Another possibility is that more automatic forms of inner speech may rely more importantly on associate memory-based processes whereas less automatic (more intentional or *deliberate*) forms of inner speech may rely more on the simulation (or emulation) mechanism [cf. the model proposed in @loevenbruck_cognitive_2018]. Why would that be the case? Automatic forms of inner speech (e.g., poem recitation)--but also more general forms of motor *habits*-- are developed through the repeated learning of the association between motor commands and the sensory consequences of these motor commands. Through repeated learning, these actions become *automatic*. By automatic, we mean that these actions i) can be executed without awareness of the action being executed, ii) can be initiated without awareness or deliberate attention, iii) can be evoked automatically by stimuli in the environment, without deliberately orienting the attention to it, and iv) are said to be automatic if they can be performed without interfering with other tasks [@norman_attention_1986]. In contrast, we may speculate that novel (unusual) motor actions, to be imagined, need to go through the simulation/emulation mechanism. This idea could be tested experimentally by creating habits (via learning) of different words and comparing their EMG traces or by assessing their "suppressibility" by articulatory suppression [see for instance @saeki_breaking_2013]. In other words, the motor imagery (or inner speech) of novel versus known material would be underpinned by different processes that would involve the motor system to a different extent. This distinction echoes @pickering_integrated_2013's distinction between the prediction-by-association and prediction-by-simulation mechanisms in language perception and comprehension. They suggested that the prediction-by-association mechanism relies more on perceptual experiences and domain-general cognitive abilities (such as memory) whereas the prediction-by-simulation mechanism would rely more simulation of the motor action leading to the speech percept. These two mechanisms may be used conjointly and weighted differently according to the task to be performed. The question remains, however, to know how this weighting is performed. We may speculate that for each task to be performed, an astute test is first performed (for instance based on familiarity), in order to determine whether the action to be performed is novel or not, and whether its consequences should be retrieved from memory (or inferred via associative mechanisms) or whether they should be simulated/emulated. In the former case, no peripheral muscular activity is expected, whereas in the latter case, the speech motor system would be involved in simulating/emulating the corresponding overt action. For simulated/emulated actions, the motor consequences of partially inhibited motor commands (i.e., small residual movements) could be recorded peripherally using surface electromyography.

Moreover, as discussed in section \@ref(inner-speech-theories), there is currently a debate as to the best architecture to model the control of motor actions, and this debate could be extended to inner speech production. More precisely, @pickering_getting_2014 made a distinction between two types of architectures, differing by the place forward models play in these architectures. First, in auxiliary forward models (AFM), forward models are considered as "special-purpose prediction mechanisms implemented by additional circuitry distinct from core mechanisms of perception and action". Second, in integrated forward models (IFM), forward models "lie at the heart of all forms of perception and action" [@pickering_getting_2014]. In other words, forward models are thought to be additionnal internal models specifically developped for the purpose of emulating motor actions (AFM) or the emulation and prediction function is thought to be realised by the same mechanisms that handle the production of motor actions (IFM). Relatedly, @friston_what_2011 argued for an IFM architecture and showed how motor control can be formalised in a Bayesian predictive framework, where optimal control can be seen as (active) inference. In these models, there would be no need for an inverse model, because the inverse model can be replaced by a Bayesian inversion of the forward model. According to @friston_what_2011, "Active inference eschews the hard inverse problem by replacing optimal control signals that specify muscle movements (in an intrinsic frame) with prior beliefs about limb trajectories (in an extrinsic frame)" (p.491). In this kind of model, motor commands are replaced by top-down (proprioceptive) predictions that drive the adjustment of the motor plant (i.e., that produce movements). This idea is similar to perceptual inference in sensory cortices, where descending connections convey predictions whereas ascending connections convey prediction errors [@adams_predictions_2013]. These descending signals are themselves predictions of proprioceptive consequences and may therefore play the role of a corollary discharge (without resorting to an inverse model). This is an interesting proposal, as most of the evidence supporting the role of an efference copy during inner speech production [e.g.,@ford_electrophysiological_2004;@tian_mental_2012;@tian_mental_2010;@tian_mental_2016;@tian_imagined_2018;@whitford_neurophysiological_2017] is actually evidence for the presence of a corollary discharge (leading to sensory attenuation) more than evidence for an efference copy per se. The idea that an inverse model may not be necessary for modelling and explaining inner speech production is also found in @wilkinson_auditory_2017, who suggested that inner speech production could be modelled in a *predictive processing framework* [for an introduction, see @clark_whatever_2013]. The need for an inverse model (or not) might be assessed in several ways. For instance, @pickering_getting_2014 suggested to look for double dissociations. Indeed, if there are distinct forward and inverse models, lesions (patient or temporary lesion studies) to the forward model should disrupt the ability to correct movements online or to learn new movements, but it should not prevent movements to be executed. In contrast, in an IFM account, all these abilities should be disrupted by a lesion to the forward (generative) model. Such an empirical test would be crucial in deciding between AFM and IFM architectures and might lead to a revision of current models of motor control.

### Implication of these results for rumination theories

Condensation in rumination, look into think-aloud protocols [e.g.,@lyubomirsky_why_1999]...?

Mental habit = low control (automaticity). Habits become habits through learning and association, in the same way a poem can be learned by heart... both can be more and more strongly internalised in a similar way... How to overcome (bad) mental habits? From @hertel_memory_2004:

> "In short, the best antidote to maladaptive habits is a new set of habits—not the opposite sort of habits recommended by Pollyanna and the teachings of Norman Vincent Peale (1956) and not the habits of suppression as recommended by certain grandmothers but the habits of thought control." (page 209)

<!--

Describe model of van vugt... suractivation of neagtive memory chunks ? Alternatively, rumination can be seen in a Bayesian framework as overdetermined priors... or lack of flexibility.

-->

See Fitzgerald et al. (2014): how habits are created ? Through model comparison and model averaging, simpler model (habits models) are favoured over time...

From Dolan & Doyan (2014): Model-free control is computationally efficient, since it replaces computation (i.e., the burdensome simulation of future states) with memory (i.e., stored discounted values of expected future reward); however, the forward-looking nature of the prediction error makes it statistically inefficient (Daw et al., 2005)...

## Limitations and ways forward

...

<!--

Différence entre petit effet homogène (dominant) et > effet hétéogène, réf aux travaux de haaf et rouder, further work should probably explore the domnancy of the effect

WEIRD samples...

-->

## Conclusion

...
